---
title: "Frequency of Internet Use and Social Media Engagement "
description: |
  Does the frequency of overall internet usage affect how many social media platforms participants engage with?
author:
  - name: Cianna Bedford-Petersen
    url: {}
date: 03-17-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(here)
here("data")
library(rio)
library(tidyverse)
library(colorblindr)


```

Participants were asked both how frequently they use the internet and which social media platforms they engage with. I am curious to know if those who visit the internet more frequently are also engaging in a higher number of platforms.

This data set required a fair amount of wrangling before visualization. Participants were asked to give a yes or no response regarding their engagement with eight separate social media platforms. From these eight response columns, I created one variable that totaled the number of platforms each participant engaged with. 


```{r plot 3 first iteration, fig.height=25, fig.width=30}

# Does the frequency of overall internet usage affect how many social media platforms participants engage with

pew <- import(here("data", "core_trends_survey.csv"))

# Tidy data
pew_tidy <- pew %>% 
  rename("Twitter" = "web1a", "Instagram" = "web1b", 
         "Facebook" = "web1c", "Snapchat" = "web1d", 
         "YouTube" = "web1e","WhatsApp" = "web1f", 
         "Pinterest" = "web1g", "LinkedIn" = "web1h") %>% 
  gather(platform, response, 
         Twitter , Instagram, Facebook, Snapchat, YouTube, WhatsApp, Pinterest, LinkedIn) %>% 
  select(respid, intfreq, platform, response) 

# Recode variable to have category names
pew_tidy$intfreq <-as.factor(recode(pew_tidy$intfreq, 
                                    "1" = "Almost Constantly", "2" = "Several Times A Day", 
                                    "3" = "About Once A Day", "4" = "Several Times A Week", 
                                    "5" = "< Several Times A Week", "8" = "Don't Know",
                                    "9" = "Refused to Answer"))

# Put internet frequency in an interpretable order
pew_tidy$intfreq <- factor(pew_tidy$intfreq, 
                           levels=c("Almost Constantly", "Several Times A Day", 
                                    "Several Times A Week", "< Several Times A Week", 
                                    "About Once A Day"))

# Filter to include just yes responses for each platform and exclude don't know and NA responses for internet use frequency 
pew_yes <- pew_tidy %>% 
  filter(response==1, 
         intfreq!="Don't Know", 
         intfreq!="Refused to Answer")

# Summarize data to get a total count of platorms engaged with for each participant then summarize data for each frequency level and add proportion column
prop <- pew_yes %>% 
  group_by(respid, intfreq) %>% 
  summarize(total_platform=n()) %>% 
  group_by(intfreq, total_platform) %>% 
  summarize(n = n()) %>% 
  mutate(proportion = n / sum(n))

# Make number of platforms into a categorical variable
prop$total_platform <- factor(prop$total_platform)
  
# Plot by frequency group
ggplot(prop, aes(total_platform, proportion)) + 
  geom_col() + 
  coord_flip() +
  facet_wrap(~intfreq) +
  scale_y_continuous(labels = scales::percent)+
  theme_minimal() +
  theme(axis.text.y = element_text(size = 40),
        axis.text.x = element_text(size = 35), 
        axis.title.x = element_text(size = 50), 
        axis.title.y = element_text(size = 50),
        strip.text.x = element_text(size = 40))

```

While this first plot conveys all of the necessary information, it is difficult to compare across frequency categories. To resolve this, I will put all of the facets in one row and add a fill by number of platforms engaged with. I will also add a title and clean up the labels to better orient the reader. 

```{r plot 2 second iteration, fig.height=20, fig.width=40}

# Add titles and put all facets in one row
ggplot(prop, aes(total_platform, proportion, fill = total_platform)) + 
  geom_col() + 
  coord_flip() +
  facet_wrap(~intfreq, nrow = 1) +
  scale_y_continuous(breaks = c(0,.15,.30),
                     labels = c("0%","15%","30%")) +
  labs(title = "Number of Platforms Used By Frequency of Internet Usage\n", 
       y = "\nPercentage of Users", 
       x ="Number of Platforms\n") +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        legend.position="none", 
        axis.text.y = element_text(size = 65),
        axis.text.x = element_text(size = 35), 
        plot.title = element_text(size = 90),
        axis.title.x = element_text(size = 60), 
        axis.title.y = element_text(size = 60),
        strip.text.x = element_text(size = 40))

```

Putting all of the facets on one row made the comparison across frequency groups much clearer. However, filling by the number of platforms engaged with is actually a bit distracting and confusing. The percentage of users is a proportion that adds up to 100% within a facet rather than across the facets. Therefore, it actually makes more sense to fill by the frequency of internet usage. This will direct the viewers attention to the distribution within a facet and they can more easily compare the distributions across levels of internet usage frequency. Finally, I will us a color palette that is more colorblind friendly than the default. 


```{r plot third iteration, fig.height=20, fig.width=40}

# Add fill by facet to focus the viewers attention to each compare values within a facet
ggplot(prop, aes(total_platform, proportion, fill = intfreq)) + 
  geom_col() + 
  coord_flip() +
  facet_wrap(~intfreq, nrow=1) +
  scale_y_continuous(breaks = c(0,.15,.30),
                     labels = c("0%","15%","30%")) +
  labs(title= "Number of Platforms Used By Frequency of Internet Usage\n", 
       y = "\nPercentage of Users", 
       x ="Number of Platforms\n") +
  theme_minimal() +
  theme(legend.title = element_blank(), 
        legend.position ="none", 
        axis.text.y = element_text(size = 65), 
        axis.text.x = element_text(size = 35), 
        plot.title = element_text(size = 90), 
        axis.title.x = element_text(size = 60), 
        axis.title.y = element_text(size = 60), 
        strip.text.x = element_text(size = 40)) +
  scale_fill_viridis_d()

```

The final plot clearly shows that people who infrequently use the internet are likely to engage with only one or two social media platforms, while those who use the internet more frequently are likely to engage with three to five different social media platforms. 