---
title: "Adult Attidutes Towards the Internet"
description: |
  Overall, would you consider the internet to be a good thing or a bad thing for society?
author:
  - name: Cianna Bedford-Petersen
    url: {}
date: 03-16-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(here)
here("_posts")
library(rio)
library(tidyverse)
library(colorblindr)
library(ggridges)

```

Participants were asked if overall, they consider the internet to be a good or bad thing for society. I am curious if the distribution of ages varies by response. I would predict that older participants are more likely to say that the internet is a bad thing for society than younger participants.

For the first iteration of this plot, I used ggridges to to get a sense of the age distribution for each of the possible responses.  

```{r plot 1 fist iteration}

# Import data
pew <- import(here("_posts", "core_trends_survey.csv"))


# Change response variable to a factor and give categorical response names
pew$pial11<- as.factor(recode(pew$pial11, 
                               "1" = "Good Thing", "2" = "Bad Thing", 
                               "3" = "Some of Both", "8" = "Don't Know", 
                               "9" = "Refused to Answer"))

# Change name of response variable to be more interpretable
pew <- pew %>% 
  rename(response_category = pial11 )

# Ridgeline density plot
ggplot(pew, aes(age, response_category)) +
  geom_density_ridges(color = "white",
                                alpha = 0.4,
                                fill = "#003F9C")


```

From this first plot it looks like there may be some differences in response by age but it is difficult to see what the pattern might be. I was curious to know how many respondants were in each category. In the table below you can see that only 10 participants refused to answer, this is likely not enuough to represent the population well and will be removed from future plots.

```{r plot 1 count table}

# How many are in each response category?
response_count <- pew %>% 
  group_by(response_category) %>% 
  summarize(response_count=n())
knitr::kable(response_count)

# There are only 10 in the refused to answer category so it is not a good representation to visualize
pew2 <- pew %>% 
  filter(response_category!="Refused to Answer")
```

For the second interation of the plot, I put the reponse categories in an order that was more interpretable to aide in spotting any patterns in the data. Additonally, I chose a larger bandwith to show a clearer peak for each distribution. Finally, I cleaned up titles and axes to help orient the reader.

```{r plot 1 second interation}

# Put responses in an order that makes more sense to read
pew2$response_category <- factor(pew2$response_category, 
                      levels = c("Don't Know", "Bad Thing", "Some of Both", "Good Thing"))

# Chose a larger bandwith to show a clearer peak for each distribution and clean up titles and axes

gr <- ggplot(pew2, aes(age, response_category)) +
  geom_density_ridges(color = "white",
                                alpha = 0.4,
                                fill = "#003F9C",
                                bandwidth = 8) +
  labs(title = "Adult Attitudes Towards the Internet", 
       subtitle = "Would you say the internet has mostly been a...",
       x = "Age") +
  scale_x_continuous(breaks = c(0,20,40,60,80,100), labels = c(0,20,40,60,80,100))+
  theme_minimal()+
  theme(axis.title.y = element_blank())
gr

```

At this point, our plot would be pretty interpretable to any scientific audience that was familiar with denisity distributions. However, to make this plot more approachable for a general audience, it would be helpful to have a reference point that makes it easy to compare response groups. For the final interation of the plot, I added a line and text at the highest density point for each response category. 

``` {r third iteration}

# Add a line at the highest density point for each response category

# Extract the data ggplot used to prepare the figure.
ingredients <- ggplot_build(gr) %>% purrr::pluck("data", 1)

# Pick the highest point
density_lines <- ingredients %>%
  group_by(group) %>% filter(density == max(density)) %>% ungroup()

# Add lines and text for highest points to plot
ggplot(pew2, aes(age, response_category)) +
  geom_density_ridges(color = "white",
                                alpha = 0.4,
                                fill = "#003F9C",
                                bandwidth = 8) +
  geom_segment(data = density_lines, 
               aes(x = x, 
                   y = ymin, xend = x, 
                   yend = ymin + density*scale*iscale)) +
  geom_text(data = density_lines, 
            aes(x = x, 
                y = ymin + 0.5 *(density*scale*iscale),
                label = round(x, 0)),
            hjust = -0.4, vjust = .03)+
  labs(title = "Adult Attitudes Towards the Internet", 
       subtitle = "Would you say the internet has mostly been a...",
       x = "Age") +
  scale_x_continuous(breaks = c(0,20,40,60,80,100,100), labels = c(0,20,40,60,80,100,100))+
  theme_minimal()+
  theme(axis.title.y = element_blank())

```

From the final plot we can see that older participants were slightly more likely to say that the internet has had at least some bad influence rather than being completely a good thing, although, the effect of age is not as dramatic as I originally expected. However, there is a large skew towards older participants for the response that they don't know if the internet has been mostly good or bad. This may indicate that younger people have more definitive beliefs about the impact of the internet. 